simodels: A package for Spatial Interaction Modelling

### 2025-03-07 paper/paper.qmd
Update with flows to school

---
title: "simodels: A package for Spatial Interaction Modelling"
---

<!-- 
# Note: you need to have the relevant quarto extension installed:
quarto add quarto-journals/jss
After that put the following in the yaml:
format:
    jss-pdf:
        keep-tex: true
    jss-html: default
# author:
#   - name: Achim Zeileis
#     affiliations:
#       - name: Universität Innsbruck
#         department: Department of Statistics
#         address: Universitätsstr. 15
#         city: Innsbruck
#         country: Austria
#         postal-code: 6020
#       - Journal of Statistical Software
#     orcid: 0000-0003-0918-3766
#     email: Achim.Zeileis@R-project.org
#     url: https://www.zeileis.org/
#   - name: Second  Author
#     affiliations:
#       - Plus Affiliation
# abstract: |
#   This short article illustrates how to write a manuscript for the
#   _Journal of Statistical Software_ (JSS) using its {{< latex >}} style files.
#   Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
#   it is recommended to keep the {{< latex >}} code as simple as possible,
#   i.e., avoid inclusion of packages/commands that are not necessary.
#   For outlining the typical structure of a JSS article some brief text snippets are employed that have been inspired by @ZeileisKleiberJackman2008, discussing count data regression in [R]{.proglang}. Editorial comments and instructions are marked by vertical bars.
abstract: |
  This article introduces the [simodels]{.pkg} package for spatial interaction modelling in R. The package provides a flexible framework for estimating various types of spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained. [simodels]{.pkg} offers efficient implementations of traditional spatial interaction models as well as more recent innovations, supporting both 'monopartite' and 'bipartite' spatial interaction scenarios. By streamlining the process of implementing SIMs in R, the package enables researchers and practitioners to more easily explore, analyze, and predict spatial interactions in various contexts, from transportation planning to economic geography.

keywords: [spatial interaction models, R, transportation modelling, gravity models, origin-destination data]
keywords-formatted: [spatial interaction models, "[R]{.proglang}", transportation modelling, gravity models, origin-destination data]

bibliography: bibliography.bib
execute: 
  messages: false
  warning: false 
  echo: false
-->

>>>>

---
title: "simodels: A package for Spatial Interaction Modelling"
bibliography: bibliography.bib
number-sections: true
execute: 
  messages: false
  warning: false 
  echo: false
---

```{r}
#| eval: false
#| echo: false
# Compile to docx and jss-pdf formats:
quarto::quarto_render("paper/paper.qmd", output_format = "docx")
quarto::quarto_render("paper/paper.qmd", output_format = "jss-pdf")
system("gh release upload v0.1.0 paper/paper.docx --clobber")
system("gh release upload v0.1.0 paper/paper.pdf --clobber")
```

<!-- 
# Note: you need to have the relevant quarto extension installed for the JSS version:
# quarto add quarto-journals/jss
# After that put the following in the yaml:
format:
    jss-pdf:
        keep-tex: true
    jss-html: default
# author:
#   - name: Achim Zeileis
#     affiliations:
#       - name: Universität Innsbruck
#         department: Department of Statistics
#         address: Universitätsstr. 15
#         city: Innsbruck
#         country: Austria
#         postal-code: 6020
#       - Journal of Statistical Software
#     orcid: 0000-0003-0918-3766
#     email: Achim.Zeileis@R-project.org
#     url: https://www.zeileis.org/
#   - name: Second  Author
#     affiliations:
#       - Plus Affiliation
# abstract: |
#   This short article illustrates how to write a manuscript for the
#   _Journal of Statistical Software_ (JSS) using its {{< latex >}} style files.
#   Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
#   it is recommended to keep the {{< latex >}} code as simple as possible,
#   i.e., avoid inclusion of packages/commands that are not necessary.
#   For outlining the typical structure of a JSS article some brief text snippets are employed that have been inspired by @ZeileisKleiberJackman2008, discussing count data regression in [R]{.proglang}. Editorial comments and instructions are marked by vertical bars.
abstract: |
  This article introduces the [simodels]{.pkg} package for spatial interaction modelling in R. The package provides a flexible framework for estimating various types of spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained. [simodels]{.pkg} offers efficient implementations of traditional spatial interaction models as well as more recent innovations, supporting both 'monopartite' and 'bipartite' spatial interaction scenarios. By streamlining the process of implementing SIMs in R, the package enables researchers and practitioners to more easily explore, analyze, and predict spatial interactions in various contexts, from transportation planning to economic geography.

keywords: [JSS, style guide, comma-separated, not capitalized, R]
keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]
-->

## Ideas 
xx

- Conceptual diagram showing decision points when developing SIMs, e.g.:
  - Unipartite or bipartite?
  - Unconstrained, production-constrained, attraction-constrained, or doubly-constrained? Or other?
  - Calculation and parameter optimisation or regression model?
- Open access dataset?
